\documentclass[11pt]{article}

\usepackage{amsmath,amssymb,amsthm}

\usepackage{fullpage}
\usepackage{hyperref}
\usepackage{color}
\usepackage[barr]{xy}
\usepackage{todonotes}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{stmaryrd}
\usepackage{mathpartir}
\usepackage{xspace}
\newcommand{\thelang}{\textsc{Disco}\xspace}

\title{\vspace{-50px}Statement of Work for Hendrix College}
\date{\vspace{-60px}}
\begin{document}
\maketitle  

\noindent
PI: Brent Yorgey\\

\textbf{Year 1.}  Recruit a total of four -- two during the academic
year and two during the summer -- undergraduate research assistants to
aid the PI in the implementation of the \thelang{}
environment.  The PI will also help design and develop the new
discrete mathematics curriculum.

\textbf{Year 2} Recruit a total of four -- two during the academic
year and two during the summer -- undergraduate research assistants to
aid the PI in the implementation of the \thelang{}
environment.  This year will begin the evaluation of the curriculum
developed during the first year by incorporating it into the discrete
mathematics course in the Mathematics \& Computer Science Department
at Hendrix College.  In addition, the PI will help recruit two
external faculty members to pilot the new curriculum in their discrete
mathematics course.

\textbf{Year 3} Recruit a total of four -- two during the academic
year and two during the summer -- undergraduate research assistants to
aid the PI in the implementation of the \thelang{}
environment.  The PI will help recruit six external faculty members to
pilot the new curriculum in their discrete mathematics course.  In
addition, the PI will help in hosting workshops and tutorials on the
new curriculum and \thelang{} environment.
\end{document}
