\documentclass[11pt]{article}

\usepackage{amsmath,amssymb,amsthm}

\usepackage{fullpage}
\usepackage{hyperref}
\usepackage{color}
\usepackage[barr]{xy}
\usepackage{todonotes}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{stmaryrd}
\usepackage{mathpartir}

\title{\vspace{-50px}Budget Justification}
\date{\vspace{-60px}}
\begin{document}
\maketitle  

\noindent
Total Funds: $\$296,959$\\
\vspace{-20px}
\section{Justification for Year 1}
\label{sec:year_1}
Total Funds Requested for Year 1: $\$94,561$\\
\vspace{-20px}
\subsection*{1.A Senior Personnel}
\label{subsec:a_senior_personnel}

Funds Requested: $\$17,264$\\
\ \\
Justification: These funds will be used as two months summer salary
for PI Harley Eades.  Summer salaries at Augusta University are twenty
percent of the faculty members yearly salary.  The PI's yearly salary
is $\$86,320$.

% subsection a._senior_personnel (end)

\subsection*{1.B Other Personnel}
\label{subsec:personnel}
\begin{itemize}

\item \textbf{Undergraduate Students:}\\ Funds Requested:
  $\$7,680$\\ \ \\ Justification: These funds will be used as a
  stipend to fund two undergraduate students during the first academic
  year of the grant, and two summer students.  Each student will be
  paid $\$8.00$ an hour and be required to work twenty hours per week.
\end{itemize}
% subsection personnel (end)

\subsection*{1.C Fringe Benefits}
\label{subsec:c_fringe_benefits}
Funds Requested: $\$3,108$\\
\ \\
These funds are the fringe benefits associated with eighteen percent
of the total salary and wages being requested.  
% subsection c_fringe_benefits (end)

\subsection*{1.E Travel}
\label{subsec:travel}
Funds Requested: $\$11,425$\\ \ \\ Justification: The requested funds
will be used as travel support for a yearly project meeting and two
international seven day conferences.

\ \\ The funds for the yearly project meeting will be used for the PI
Harley Eades and his two summer undergraduate research assistants to
travel to Hendrix College: (Flight) $\$3,000$, (Housing) $\$1,500$,
and (Meals) $\$300$ for a total of $\$4,800$.

\ \\ The funds for travel support for two international seven day
conferences can be broken down into the amounts: (Registration) $\$1,500.00$,
(Flight) $\$2,500$, (Housing) $\$2,100$, and (Food) $\$525$ for a total of
$\$6,625$.  
% subsection travel (end)

\subsection*{1.G Other Direct Costs}
\label{subsec:other}

\begin{itemize}
\item \textbf{Subcontracts:} Funds Requested: $\$41,196$

  Justification: These funds will be used as the subaward to Hendrix
  College for the co-PI Brent Yorgey.  The amount is broken down as
  follows:
  \begin{itemize}
  \item Undergraduate students: Requested Funds: $\$9,130$.  These
    funds are for two undergraduate students during the first academic
    year each required to work ten hours a week, and two undergraduate
    students during the summer of the first year of the proposed
    project each required to work forty-hours a week.  Each student
    will be paid $\$8.00$ per hour.
  \item Co-PI salary: Requested Funds: $\$16,144$.  These funds are
    for two months summer salary. The base yearly -- on a nine month
    contract -- salary is $\$67,485$, so two months is $\$14,997$.
    Then adding in $7.65\%$ for fringe benefits or $\$1,147$ we arrive
    at the requested amount.
  \item Travel: Requested Funds: $\$6,625$.  These funds are for
    travel support for two international seven day conferences. It is
    broken down into the amounts: (Registration) $\$1,500.00$,
    (Flight) $\$2,500$, (Housing) $\$2,100$, and (Food) $\$525$.
  \item University Overhead: Requested Funds: $\$9,298$.  These funds
    are to cover the Hendrix College university overhead which is
    calculated by taking $62\%$ of the Co-PI's base salary
    ($\$14,997$) for two summer months.
  \end{itemize}

\item \textbf{Other:} Funds Requested: $\$480$\\ \ \\ Justification:
  These funds will be used to pay for cloud hosting of the web-based
  IDE that will be developed as part of the proposed project.
\end{itemize}


\section*{1.I Indirect Costs}
\label{sec:1.i_indirect_costs}
Funds Requested: $\$13,409$\\
\ \\
Justification: Augusta University charges $47.8\%$ of the
total salaries, wages and fringe benefits requested.  Thus, $47.8\%$
of $\$28,052$ is the requested amount.
% section 1.i_indirect_costs (end)
% section year_1 (end)

\section{Justification for Year 2}
\label{sec:justification_for_year_2}

Total Funds Requested for Year 2: $\$98,199$\\

\noindent
We only detail the justifications from year 2 that differ from year 1.

\subsection*{2.E Travel}
\label{subsec:travel}
Funds Requested: $\$6,625$\\ \ \\ Justification: The funds for travel
support for two international seven day conferences can be broken down
into the amounts: (Registration) $\$1,500.00$, (Flight) $\$2,500$,
(Housing) $\$2,100$, and (Food) $\$525$ for a total of $\$6,625$.
% subsection travel (end)

\subsection*{2.F Participant Support Cost }
\label{subsec:2.f_participant_support_cost_}
\begin{itemize}
\item Stipends: Requested Funds: $\$3,000$.  These funds will be used
  to support two recruited faculty members for piloting the proposed
  project in their discrete mathematics course.
\end{itemize}
% subsection 2.f_participant_support_cost_ (end)

\subsection*{2.G Other Direct Costs}
\label{subsec:other}

\begin{itemize}
\item \textbf{Subcontracts:} Funds Requested: $\$46,634$

  Justification: These funds will be used as the subaward to Hendrix
  College for the co-PI Brent Yorgey.  The amount $\$41,196$ of the
  requested amount is justified in the same way as year 1, but we add
  $\$4,800$ to support Brent and his summer students to attend the
  yearly project meeting at Augusta University, and $\$638$ of
  additional fringe benefits.
\end{itemize}

% section justification_for_year_2 (end)

\section{Justification for Year 3}
\label{sec:justification_for_year_3}

Total Funds Requested for Year 3: $\$104,199$\\

\noindent
We only detail the justifications from year 3 that differ from year 1.

\subsection*{3.F Participant Support Cost }
\label{subsec:3.f_participant_support_cost_}
\begin{itemize}
\item Stipends: Requested Funds: $\$9,000$.  These funds will be used
  to support six recruited faculty members for piloting the proposed
  project in their discrete mathematics course.
\end{itemize}

\subsection*{3.G Other Direct Costs}
\label{subsec:other}

\begin{itemize}
\item \textbf{Subcontracts:} Funds Requested: $\$41,834$

  Justification: These funds will be used as the subaward to Hendrix
  College for the co-PI Brent Yorgey.  The amount $\$41,196$ of the
  requested amount is justified in the same way as year 1, but we add
  $\$638$ of additional fringe benefits.
\end{itemize}

% section justification_for_year_3 (end)
\end{document}
